Case for Support

### 2021-11-16 application/detailsOfResource.tex
11/16 changes Maneesh
Merged Figs 2 and 3

We intend to improve an \textbf{existing resource}, the Bonsai ecosystem. We
propose to create software infrastructure to enable the\textbf{ integration of advanced data analysis methods} into
Bonsai (Section~\ref{sec:infrastructure}).
%
To demonstrate the use of this infrastructure, and to provide Bonsai users with
an initial set of advanced machine intelligence neural data analysis methods,
we will add to the Bonsai ecosystem machine learning functionality to control
neuroscience experiments and to analyse data generated by them
(Section~\ref{sec:functionality}).

@@

\begin{figure}
  \includegraphics[width=4in]{figures/proposed_bonsai_extensions.png}
  \caption{Proposed extensions to Bonsai}
  \label{fig:proposedBonsaiExtensions}
\end{figure}

@@

\begin{comment}
Bonsai does not currently provide machine learning functionality. For instance,
the close-loop experimental control depicted in
Figure~\ref{fig:bci} cannot be currently implemented in
Bonsai. It represents the control of an experiment where a subject guides the
mouse position in a computer screen with her brain activity (Section~\ref{}).
Nodes highlighted in colour require machine intelligence functionality currently
not available in Bonsai.
%
\end{comment}

@@

Inspired by \texttt{scikit-learn}, Bonsai will have three types of operators:
\texttt{Estimators}, \texttt{Predictors} and \texttt{Transformers}. Each of
these operators will be instantiated by objects implementing the corresponding
interface.
%
As an example, consider the close-loop system for the control of a computer
cursor with neural activity represented in Figure~\ref{fig:bci}a. The Bonsai
graphical program for this system is given in Figure~\ref{fig:bci}b. The
objects instantiating the SpikeThresholdDetector, the KiloSortSpikeSorter and
GPFAlatentsExtractor will implement the \texttt{Transformer} interface (blue),
and the object instantiating the TimeSeriesForestClassifier will implement the
\texttt{Predictor} interface (green; Figure~\ref{fig:bci}b, top).
%
The parameters for the SpikeThresholdDetector and the KiloSortSpikeSorter will
be set manually by the user, as is commonly done in practice. The parameters of
the GPFAlatentsExtractor will be learnt offline in an unsupervised way by an
\texttt{Estimator} (red) in a second pipeline (Figure~\ref{fig:bci}b, middle).
Finally, the parameters of the TimeSeriesForestClassifier will also be learnt
offline, but in a supervised way, by another \texttt{Estimator} in a third
pipeline (Figure~\ref{fig:bci}b, bottom).

\begin{figure}
   \begin{center}
       \includegraphics[width=4in]{figures/bci.png}
       \caption{Close-loop system for controlling a computer cursor with brain activity}
     \label{fig:bci}
   \end{center}
\end{figure}

\begin{comment}
    The spike detector transforms membrane
potentials into spike sequences. The spike sorter transforms spikes sequences
into lists of spikes, one list per neuron. The latents extractor transforms
multi-neuron lists of spikes into latent trajectories summarising the
multi-neuron spiking activity.  The neural decoder predicts the next cursor
position from the latent trajectories.
\end{comment}

@@

We propose to add functionality into Bonsai for the implementation of
brain-computer-interface (BCI) applications illustrated in Figure~\ref{fig:bci}. In
this implementation voltage recordings from the brain will be converted into
spikes fired by single neurons by a \texttt{SpikeDetector}
(Section~\ref{sec:spikeDetection}), or to local field potentials (LFP) by an
\texttt{LFPextractor} (Section~\ref{sec:lfpExtraction}). Next, neural activity
(spikes or LFPs) will be represented in a low-dimensional latent space by a
\texttt{LatentsCalculator}. Finally, a \texttt{Decoder} will extract the
intended behavior of the animal from this low-dimensional representation.

>>>>

We propose to create software infrastructure to enable the\textbf{ integration of advanced data analysis methods} into
Bonsai (Section~\ref{sec:infrastructure}).
%
To demonstrate the use of this infrastructure, and to provide Bonsai users with
an initial set of advanced machine intelligence neural data analysis methods,
we will add to the Bonsai ecosystem machine learning functionality to control
neuroscience experiments and to analyse data generated by them
(Section~\ref{sec:functionality}).

@@

\begin{figure}
    \begin{center}
        \includegraphics[width=5in]{figures/proposed_bonsai_extensions_combined.png}
        \caption{Proposed extensions to Bonsai}
        \label{fig:proposedBonsaiExtensions}
    \end{center}
\end{figure}

@@

\begin{comment}
Bonsai does not currently provide machine learning functionality. For instance,
the close-loop experimental control depicted in
Figure~\ref{fig:proposedBonsaiExtensions}b cannot be currently implemented in
Bonsai. It represents the control of an experiment where a subject guides the
mouse position in a computer screen with her brain activity (Section~\ref{}).
Nodes highlighted in colour require machine intelligence functionality currently
not available in Bonsai.
%
\end{comment}

@@

Inspired by \texttt{scikit-learn}, Bonsai will have three types of operators:
\texttt{Estimators}, \texttt{Predictors} and \texttt{Transformers}. Each of
these operators will be instantiated by objects implementing the corresponding
interface.
%
As an example, consider the close-loop system for the control of a computer
cursor with neural activity represented in
Figure~\ref{fig:proposedBonsaiExtensions}b. The Bonsai
graphical program for this system is given in
Figure~\ref{fig:proposedBonsaiExtensions}c. The
objects instantiating the SpikeThresholdDetector, the KiloSortSpikeSorter and
GPFAlatentsExtractor will implement the \texttt{Transformer} interface (blue),
and the object instantiating the TimeSeriesForestClassifier will implement the
\texttt{Predictor} interface (green; Figure~\ref{fig:proposedBonsaiExtensions}c, top).
%
The parameters for the SpikeThresholdDetector and the KiloSortSpikeSorter will
be set manually by the user, as is commonly done in practice. The parameters of
the GPFAlatentsExtractor will be learnt offline in an unsupervised way by an
\texttt{Estimator} (red) in a second pipeline
(Figure~\ref{fig:proposedBonsaiExtensions}c, middle).
Finally, the parameters of the TimeSeriesForestClassifier will also be learnt
offline, but in a supervised way, by another \texttt{Estimator} in a third
pipeline (Figure~\ref{fig:proposedBonsaiExtensions}c, bottom).

\begin{comment}
\begin{figure}
   \begin{center}
       \includegraphics[width=4in]{figures/bci.png}
       \caption{Close-loop system for controlling a computer cursor with brain activity}
     \label{fig:bci}
   \end{center}
\end{figure}

    The spike detector transforms membrane
potentials into spike sequences. The spike sorter transforms spikes sequences
into lists of spikes, one list per neuron. The latents extractor transforms
multi-neuron lists of spikes into latent trajectories summarising the
multi-neuron spiking activity.  The neural decoder predicts the next cursor
position from the latent trajectories.
\end{comment}

@@

We propose to add functionality into Bonsai for the implementation of
brain-computer-interface (BCI) applications illustrated in
Figure~\ref{fig:proposedBonsaiExtensions}b. In
this implementation voltage recordings from the brain will be converted into
spikes fired by single neurons by a \texttt{SpikeDetector}
(Section~\ref{sec:spikeDetection}), or to local field potentials (LFP) by an
\texttt{LFPextractor} (Section~\ref{sec:lfpExtraction}). Next, neural activity
(spikes or LFPs) will be represented in a low-dimensional latent space by a
\texttt{LatentsCalculator}. Finally, a \texttt{Decoder} will extract the
intended behavior of the animal from this low-dimensional representation.
